
---
title: "R Notebook For DBMS Practicum 1"
output: html_notebook
Authors: Harshita Yeddula
---

#Importing the Libraries and connecting to the remote server (AWS)
```{r}
library(dplyr)
library(RMySQL)
library(sqldf)
library(dbplyr)

con <- DBI::dbConnect(RMySQL::MySQL(), 
  dbname = "dbmspracticum",
  host = "dbmspracticum.clx781tgbh1l.us-east-1.rds.amazonaws.com",
  user = "admin",
  password = rstudioapi::askForPassword("admin123")
)

```

#Creating required tables, airports, conditions and incidents with necessary primary keys and foreign keys.



#Question 1 and Question 2 - Creating tables and loading csv data into data frame

#Dropping table incidents if Exists
```{sql connection=con}
DROP TABLE IF EXISTS incidents
```

#Dropping table airports if Exists
```{sql connection=con}
DROP TABLE IF EXISTS airports
```

#Dropping table conditions if Exists
```{sql connection=con}
DROP TABLE IF EXISTS conditions
```


#Creating table conditions.
```{sql connection=con}

create table conditions(
cid INTEGER NOT NULL primary key AUTO_INCREMENT,
cond VARCHAR(200) NOT NULL,
explanation VARCHAR(200) NOT NULL
);

```


#Creating table airports
```{sql connection=con}

create table airports(
aid INTEGER NOT NULL primary key AUTO_INCREMENT,
airportName VARCHAR(200) NOT NULL,
airportCode VARCHAR(200), 
state VARCHAR(200)
);

```


#Creating table incidents
```{sql connection=con}

CREATE Table incidents(
iid INT NOT NULL AUTO_INCREMENT primary key,
date DATE NOT NULL,
origin INTEGER NOT NULL,
airline VARCHAR(200) NOT NULL, 
aircraft VARCHAR(200) NOT NULL,
flightPhase VARCHAR(200) NOT NULL,
impact BOOLEAN,
cond INTEGER NOT NULL
)

```


#Altering the table by setting the Foreign key
```{sql connection=con}
ALTER TABLE incidents
ADD FOREIGN KEY (origin) REFERENCES airports(aid);

```


#Altering the table by setting the Foreign key
```{sql connection=con}
ALTER TABLE incidents
ADD FOREIGN KEY (cond) REFERENCES conditions(cid);
```



#Loading the csv file into the Data Frame.
```{r loadCSV}
fn = "BirdStrikesData.csv"

getwd()
df.raw <- read.csv(file = fn,
                   header = T,
                   stringsAsFactors = F)
```




#Performing tasks that are required to load the tables like harmonizing the flight phases, removing the Military flights.
```{r}
library(tidyverse)
library(lubridate)


# Getting the data from the data frame that is required to insert into tables.
df.need <- data.frame(date = df.raw$FlightDate, 
                      airline = df.raw$Aircraft..Airline.Operator,
                      aircraft = df.raw$Aircraft..Make.Model,
                      flightPhase = df.raw$When..Phase.of.flight,
                      impact = df.raw$Effect..Indicated.Damage,
                      airportName = df.raw$Airport..Name,
                      state = df.raw$Origin.State,
                      condition = df.raw$Conditions..Sky)

#Linking the airportName that is NULL with sentinel
df.need$airline[which(df.need$airline =='')] <- 'sentinel'
df.need$airportName[which(df.need$airportName == '')] <- 'sentinel'


# Updating the impact as TRUE or FALSE based on damage caused or not for the impact field.
df.need$impact <- ifelse(df.need$impact == 'Caused damage', 'TRUE', 'FALSE')


# Harmonizing the flight phases.
df.need$flightPhase[which(df.need$flightPhase == 'Approach' | 
                            df.need$flightPhase == 'Landing Roll')] <- "landing"
df.need$flightPhase[which(df.need$flightPhase == 'Climb' |
                            df.need$flightPhase == 'Take-off run')] <- "takeoff"
df.need$flightPhase[which(df.need$flightPhase == 'Descent')] <- "inflight"
df.need$flightPhase[which(df.need$flightPhase == 'Taxi' | 
                            df.need$flightPhase == 'Parked' | 
                            df.need$flightPhase == '')] <- "unknown"
df.need$date <- as.Date(mdy_hm(df.need$date))


#Removing the incidents that has airline as MILTARY
df.need$airline[df.need$airline == 'MILITARY'] <- NA
df.omit <- df.need[!is.na(df.need$airline),]
```




#Creating the actual dataframes to insert into the tables.
```{r}
# Creating data frame for incidents table.
df.incidents <- data.frame(iid = seq(1, nrow(df.omit)),
                           date = df.omit$date,
                           origin = 1,
                           airline = df.omit$airline,
                           aircraft = df.omit$aircraft,
                           flightPhase = df.omit$flightPhase,
                           impact = df.omit$impact,
                           cond = 1)


# Creating data frame for airports table.
df.airportsR <- data.frame(airportName = df.omit$airportName,
                          airportCode = "",
                          state = df.omit$state)
df.airportsR <- unique(df.airportsR[, 1:3])
df.airports <- data.frame(aid = seq(1, nrow(df.airportsR)),
                          airportName = df.airportsR$airportName,
                          airportCode = "",
                          state = df.airportsR$state)

# Creating data frame for conditions table.
df.conditionsR <- data.frame(condition = df.omit$condition,
                             explanation = "")
df.conditionsR <- unique(df.conditionsR[1])
df.conditions <- data.frame(cid = seq(1, nrow(df.conditionsR)),
                            condition = df.conditionsR$condition,
                            explanation = "")
```


# Foreign keys Linking
```{r}
# Linking the foreign keys from incidents table for aid
for (r in 1:nrow(df.incidents)) {
  a <- df.airports$aid[which(df.airports$airportName == df.omit$airportName[r] &
                               df.airports$state == df.omit$state[r])]
  df.incidents$origin[r] <- a
}

# Linking the foreign keys from incidents table for cid
for (r in 1:nrow(df.incidents)) {
  b <- df.conditions$cid[which(df.conditions$condition == df.omit$condition[r])]
  df.incidents$cond[r] <- b
}
```




#Writing data into the incidents table.
```{r}
dbWriteTable(con, "incidents", df.incidents, overwrite = T, row.names = FALSE)
```


#Writing data into the airports table.
```{r}

dbWriteTable(con, "airports", df.airports, overwrite = T, row.names = FALSE)
```


#Writing data into the conditions table.
```{r}
dbWriteTable(con, "conditions", df.conditions, overwrite = T, row.names = FALSE)
```



#Question 3

#Displaying parts of the data using R.
```{r}


dbGetQuery(con, "SELECT * from incidents Limit 20")

dbGetQuery(con, "SELECT * from airports Limit 20")

dbGetQuery(con, "SELECT * from conditions Limit 20")

```



#Question 4

#Displayig number of bird strikes for each flight phase.
```{sql connection=con}
SELECT DISTINCT(flightPhase), count(iid) FROM incidents
GROUP BY flightPhase;
```


#Question 5
# Displaying FLight phases that has above average number of bird strike incidents.
```{sql connection=con}
SELECT DISTINCT(flightPhase), imp FROM (
  SELECT DISTINCT(flightPhase), count(iid) as imp FROM incidents
  GROUP BY flightPhase
)as avg
WHERE imp > (SELECT COUNT(iid)/count(distinct(flightPhase)) from incidents)
```


#Question 6
# Displaying the average number of bird strike incidents per month.
```{sql connection=con}
SELECT DISTINCT(month(date)) as month, count(iid) as NumberOfIncidents, count(iid)/12 as AverageBirdStrikeIncidents FROM incidents
GROUP BY month ORDER BY month ASC
```



#Question 7
#In order to plot the graph I am jus displaying the number of bird strike incidents int he years 2005 to 2011.
```{sql connection=con}
select COUNT(iid), YEAR(date) as year from dbmspracticum.incidents where YEAR(date) >= 2005 and YEAR(date) <= 2011 group by year Order by year desc;

```



#Question 7
#Showing the plot by displaying the year in x axis and number of bird strike incidents in that year in y axis.
```{r}

sqlCmd = "select COUNT(iid) as count, YEAR(date) as year from dbmspracticum.incidents 
 group by year
 having year >=2005 and year<=2011
 Order by year desc;";

rs2 = dbGetQuery(con, sqlCmd)


plot(x=rs2$year, y=rs2$count, xlab ="Year", ylab ="Number of Bird Strikes", main="Bird Strikes by Year")

lines(rs2$year, rs2$count, pch=18, col="blue", type="b", lty=2)

legend(2005, 2011, legend=c("Line 1" ),
       col=c("blue"), lty=1:2)
```

#Question 8
#Dropping the procedure if exists.
```{sql connection=con}
DROP PROCEDURE IF EXISTS NewBirdStrikeIncident;
```



#Creating the procedure to insert a new data.
```{sql connection=con}
create procedure NewBirdStrikeIncident(
	  IN	i_date				TEXT,
    IN	i_airline			TEXT,
    IN	i_aircraft			TEXT,
    IN	i_flightPhase		TEXT,
    IN	i_impact			TEXT,
    IN	a_airportName		TEXT,
    IN	a_state				TEXT,
    IN	c_cond				TEXT)
BEGIN
  DROP TABLE IF EXISTS NBS;
	CREATE TABLE NBS(
	did INT NOT NULL,
	date TEXT NOT NULL,
	airline TEXT,
	aircraft TEXT,
	flightPhase TEXT,
	impact TEXT,
	airportName TEXT,
	state TEXT,
	cond TEXT,
	PRIMARY KEY (did)
	);
	INSERT INTO NBS(did, date, airline, aircraft, flightPhase, impact, airportName, state, cond)
	VALUES (did, i_date, i_airline, i_aircraft, i_flightPhase, i_impact, a_airportName, a_state, c_cond);
	
INSERT INTO airports ( airportName, state) 
  SELECT airportName, state FROM NBS 
  WHERE NOT EXISTS (SELECT * FROM airports 
    WHERE airports.airportName = NBS.airportName);

INSERT INTO incidents (date, origin, airline, aircraft, flightPhase, impact, cond)
  SELECT date, airports.aid, airline, aircraft, flightPhase, impact, conditions.cid FROM NBS
  INNER JOIN conditions ON NBS.cond = conditions.condition
  INNER JOIN airports ON NBS.airportName = airports.airportName;
  
END
```



# Use Procedure NewBirdStrikeIncident to insert a new record into the database.
```{sql connection=con}
CALL NewBirdStrikeIncident ('2021-09-06','NEW AIRPORT','NEW-B-789','landing','TRUE','new','MA','No Cloud');
    
```


# Check that the data is imported correctly into Table incidents.
```{sql connection=con}
select * from incidents where airline = "NEW AIRPORT"
```

