---
title: "R Notebook For DBMS Practicum 2 - Part 2"
output: html_notebook
Authors: HARSHITHA YEDDULA 
Email: yeddula.h@northeastern.edu
---

# Task 2 Part 2 - Creating a MySQL database using a cloud MySQL instance and also connecting to local SQlite DB created in Task 1
```{r}
library(RMySQL)
library(DBI)
library(ggplot2)

conn <-  DBI::dbConnect(RMySQL::MySQL(),
  user = "admin",
  password = rstudioapi::askForPassword("admin123"),
  host = "practicum.ce5djxzipyp7.us-east-1.rds.amazonaws.com",
  port = 3306,
  dbname = "practicum"
)
```


```{r}
dbpath <- "C:/DBMS/SQL"  #Path of database
dbfile <- "pubmed.db"                 #Name of database

db <- dbConnect(RSQLite::SQLite(), paste0(dbpath, dbfile))
```

```{r}
dbListTables(db)
```


#Task 2 Part3 - Creating star schema

#Dropping Table AuthorFact table
```{sql connection=conn}
DROP TABLE IF EXISTS AuthorFact;
```

#Creating AuthorFact table
```{sql connection = conn}
CREATE TABLE AuthorFact
( 
AuthorId INTEGER PRIMARY KEY,
LastName TEXT,
FirstName TEXT,
NumArticle INTEGER ,
AvgArticle NUMERIC);
```

#Dropping Table JournalsDimension table
```{sql connection=conn}
DROP TABLE IF EXISTS JournalsDimension;
```

#Creating JournalsDimension table
```{sql connection = conn}
CREATE TABLE JournalsDimension
( 
Issue_Id INTEGER PRIMARY KEY,
Title TEXT,
Issue INTEGER,
Volume INTEGER,
PubDate date);
```

#Dropping Table ArticlesDimension table
```{sql connection=conn}
DROP TABLE IF EXISTS ArticlesDimension;
```

#Creating ArticlesDimension table
```{sql connection = conn}
CREATE TABLE ArticlesDimension
( 
ArticleId INTEGER PRIMARY KEY,
ArticleTitle TEXT,
Year DATE, 
Month DATE);
```

#Dropping Table AuthorshipDimension table
```{sql connection=conn}
DROP TABLE IF EXISTS AuthorshipDimension;
```

#Creating AuthorshipDimension table
```{sql connection = conn}
CREATE TABLE AuthorshipDimension
(
AuthorId INTEGER PRIMARY KEY,
ArticleId INTEGER NOT NULL,
CONSTRAINT has FOREIGN KEY (ArticleId) REFERENCES ArticlesDimension (ArticleId),
CONSTRAINT FOREIGN KEY (AuthorId) REFERENCES AuthorFact (AuthorId)
);
```

#Dropping Table JournalOwnershipDimension table
```{sql connection=conn}
DROP TABLE IF EXISTS JournalOwnershipDimension;
```

#Creating JournalOwnershipDimension table
```{sql connection = conn}
CREATE TABLE JournalOwnershipDimension
(
ArticleId INTEGER NOT NULL,
Issue_Id INTEGER NOT NULL,
CONSTRAINT PK_Journal_Ownership PRIMARY KEY (ArticleId,Issue_Id), CONSTRAINT  FOREIGN KEY (ArticleId) REFERENCES ArticlesDimension (ArticleId),
CONSTRAINT FOREIGN KEY (Issue_Id) REFERENCES JournalsDimension (Issue_Id) 
);
```

#Reading tables from SQLite database and populating the data frames.
```{r}
Article_df =  dbReadTable(db, "Articles")
Authors_df =  dbReadTable(db, "Authors")
Journals_df =  dbReadTable(db, "Journals")
Authorship_df =  dbReadTable(db, "Authorship")
Journal_Ownership_df =  dbReadTable(db, "JournalOwnership")

```

#Populating tables using the data frames.
```{r}
dbWriteTable(conn,"Articles",Article_df , row.names=F,append=T)
dbWriteTable(conn,"Authors",Authors_df , row.names=F,append=T)
dbWriteTable(conn,"Journals",Journals_df , row.names=F,append=T)
dbWriteTable(conn,"Authorship",Authorship_df , row.names=F,append=T)
dbWriteTable(conn,"JournalOwnership",Journal_Ownership_df , row.names=F,append=T)

```

#Querying on AuthorsFact table
```{r}
Authors_Fact<-dbGetQuery(conn,"select AuthorId,LastName, FirstName,count(ArticleId) as NumArticle, ROUND(count(*)/count(distinct year),0) as AvgArticle
from
(select au.AuthorId,au.LastName, au.FirstName,Title,jo.Issue_Id,aus.ArticleId,ar.Year
from 
Articles AS ar inner join Authorship  AS aus  On aus.ArticleId =ar.ArticleId

inner join Authors AS au On au.AuthorId=aus.AuthorId

inner join JournalOwnership AS jo ON jo.ArticleId =aus.ArticleId
inner join Journals AS j On j.Issue_Id =jo.Issue_Id

Order by au.AuthorId) tb1
group by AuthorId")
```

#Writing to table AuthorsFact
```{r}
dbWriteTable(conn,"AuthorsFact",Authors_Fact, row.names=F,append=T)
```

```{r}
dbWriteTable(conn=conn,"AriclesDimension",value=Article_df, append= TRUE,row.names=FALSE)
dbWriteTable(conn=conn,"AuthorsFact",value=Authors_Fact, append= TRUE,row.names=FALSE)
dbWriteTable(conn=conn,"JournalsDimension",value=Journals_df , append= TRUE,row.names=FALSE)
dbWriteTable(conn=conn,"AuthorshipDimension",value=Authorship_df ,append= TRUE,row.names=FALSE)
dbWriteTable(conn=conn,"JournalOwnershipDimension",value=Journal_Ownership_df,append= TRUE,row.names=FALSE)
```

# Task 3 - queries using your MySQL data warehouse to populate a fictitious dashboard 

#Getting Top 10 data
```{r}
Top10 <- dbGetQuery(conn,"select*
from AuthorsFact
order by NumArticle desc limit 10")
```


#Populating the Top 10 data
```{r}
ggplot(data = Top10, aes(x = FirstName, y=  NumArticle)) + geom_bar(position = 'dodge', stat = 'identity') + ggtitle("Top 10 Author's publications")
```


```{r}
seasonal <- dbGetQuery(conn,"SELECT a.Quarter, COUNT(a.ArticleID) AS NumArticles
FROM( SELECT Month,ArticleID,
CASE 
     WHEN Month = 1 THEN 'Quarter_1'
     WHEN Month = 2 THEN 'Quarter_1'
     WHEN Month = 3 THEN 'Quarter_1'
     WHEN Month = 4 THEN 'Quarter_2'
     WHEN Month = 5 THEN 'Quarter_2'
     WHEN Month = 6 THEN 'Quarter_2'
     WHEN Month = 7 THEN 'Quarter_3'
     WHEN Month = 8 THEN 'Quarter_3'
     WHEN Month = 9 THEN 'Quarter_3'
     WHEN Month = 10 THEN 'Quarter_4'
     WHEN Month = 11 THEN 'Quarter_4'
     WHEN Month = 12 THEN 'Quarter_4'
END AS Quarter
FROM Articles) a 
GROUP BY Quarter")
```


```{r}
ggplot(data = seasonal, aes(x = Quarter, y= NumArticles )) + geom_bar(position = 'dodge', stat = 'identity') + ggtitle("The Number of publications for each Quarter")
```

#Disconneting the connections
```{r}
dbDisconnect(conn)
dbDisconnect(db)
```